Open Clusters Characterization in Gaia DR2 Using ML Algorithms

### 2020-11-11 thesis/thesis.tex
thesis: Add Soft Clustering with K-Means section

\usepackage{caption}
\usepackage{eso-pic}
\usepackage{float}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{picture}
\usepackage{subcaption}

@@

@@

@@

% TODO: Hablar de cómo funciona el descargador y cómo está formada la base de datos

@@

% TODO: Describir cómo funciona el clustering con KMedias

@@

% TODO: Describir cómo funciona el clustering usando dec

>>>>

\usepackage{apacite}
\usepackage{caption}
\usepackage{eso-pic}
\usepackage{float}
\usepackage{graphicx}
\usepackage{picture}
\usepackage{subcaption}

\hypersetup{
  colorlinks = true, % Colors links instead of ugly boxes
  urlcolor   = azulunir, % Color for external hyperlinks
  linkcolor  = azulunir, % Color of internal links
  citecolor  = azulunir, % Color of citations
}

@@

% TODO: Hablar sobre los algoritmos de clustering que se van a emplear: KMeans, DEC

@@

Methods for retrieving information from the database are available through an instance of \verb|cdalvaro.DB|. This class contains methods
for recovering stars information for a given region as a Pandas \verb|DataFrame| ready for data analysis.

@@

@@

Our first approach to find the open cluster inside the interest region is to use the K-Means algorithm.

Since we are looking for a cluster, it seems reasonable to use a clustering algorithm and set it to find two clusters.
One for the desired OC and another cluster which contains stars that don't belong to the open cluster.
However, this idea is not right.

This is due to the fact that OC's stars are sourounded by other stars with possibly similar properties. So, setting the number of clusters
to two is too low to separate them properly.

As shown in Figure \ref{fig:kmeans_comparisons_melotte_22}, bigger values for the number of clusters allow us to isolate more accurately
the resonance in parallax at $\approx 7.5mas$. However, we have the disadvantage that more groups are formed, while our aim is to find just
one small cluster and another with the rest of stars. This effect complicates our task of finding the desired OC. Thus, we have to find
the right value to isolate the searched cluster without creating to many groups.

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n2_pm_melotte_22.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n5_pm_melotte_22.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n8_pm_melotte_22.pdf}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n2_parallax_melotte_22.pdf}
      \caption{N clusters = 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n5_parallax_melotte_22.pdf}
      \caption{N clusters = 5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans_n8_parallax_melotte_22.pdf}
      \caption{N clusters = 8}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means comparisons}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure}

To solve this issue, we will try to estimate the best number of clusters by using the \emph{silhouette score}.

The silhouette score is a metric used to determine how good a cluster is based on intra-cluster distances and nearest-cluster distances.
The best possible value is 1, and the worst is -1. Negative values mean that some samples have been assigned to the wrong cluster.

@@

Our initial approach allows us to find some groups that potentially contain the OC. However we need now a method to merge clusters
by migrating stars from one group to another. So we can end up with the minimum number of clusters while preserving the open cluster.

For that purpose we have adapted the Unsupervised Deep Embedding for Clustering Analysis or DEC model \cite{xie2016unsupervised} to our work.

DEC is divided in two stages. In the first one and \emph{deep autoencoder} is pretrained to extract information from data and generate the
the embedding vectors. In the second stage, a parameter optimization is performed, iterating between computing an auxiliary target distribution
and minimizing the Kullback-Leiber (KL) divergence \cite{kullback1951information} to it.

This model uses K-Means as the starting point, since it is used to create an initial clustering and then the DEC model is used to refined it.
